
# Inequalities, Logical expressions

## Boolean values

In mathematics it is common to test if an expression is true or false. For example, is the point $(1,2)$ inside the disc $x^2 + y^2 \leq 1$? We would check this by substituting $1$ for $x$ and $2$ for $y$, evaluating both sides of the inequality and then assessing of the relationship is true or false. In this case, we end up with a comparison of $5 \leq 1$, which we of course know is false.

`Julia` provides numeric comparisons that allow this notation to be exactly mirrored:

```{r eval=FALSE}
x, y = 1, 2
x^2 + y^2 <= 1

```
                                    

```{r eval=FALSE}
## false
```



The response is `false`, as expected. `Julia` provides [Boolean](http://en.wikipedia.org/wiki/Boolean_data_type) values `true` and `false` for such questions. The same process is followed as was described mathematically.

The set of numeric comparisons is nearly the same as the mathematical counterparts: `<`, `<=`, `==`, `>=`, `>`. The syntax for less than or equal can also be represented with the Unicode `≤` (generated by `\le[tab]`). Similarly, for greater than or equal, there is `\ge[tab]`.
<div class="alert alert-success" role="alert">
<div class="markdown"><p>The use of <code>&#61;&#61;</code> is necessary, as <code>&#61;</code> is used for assignment.</p>
</div> 
</div>

The `!` operator takes a boolean value and negates it. It uses prefix notation:

```{r eval=FALSE}
!true

```
                                    

```{r eval=FALSE}
## false
```



For  convenience, `a != b` can be used in place of `!(a == b)`.

## Algebra of inequalities

To illustrate, let's see that the algebra of expressions works as expected.

For example, if $a < b$ then for any $c$ it is also true that $a + c < b + c$.

We can't "prove" this through examples, but we can investigate it by the choice of various values of $a$, $b$, and $c$. For example:

```{r eval=FALSE}
a,b,c = 1,2,3
a < b, a + c < b + c

```
                                    

```{r eval=FALSE}
## (true,true)
```



Or in reverse:

```{r eval=FALSE}
a,b,c = 3,2,1
a < b, a + c < b + c

```
                                    

```{r eval=FALSE}
## (false,false)
```



Trying other choices will show that the two answers are either both `false` or both `true`.
<div class="alert alert-success" role="alert">
<div class="markdown"><p>Well, almost... When <code>Inf</code> or <code>NaN</code> are involved, this may not hold, for example <code>1 &#43; Inf &lt; 2 &#43; Inf</code> is actually <code>false</code>. </p>
</div> 
</div>

So adding or subtracting any finite value from an inequality will preserve the inequality, just as it does for equations.

What about addition and multiplication?

Consider the case $a < b$ and $c > 0$. Then $ca < cb$. Here we investigate using 3 random values (which will be positive):

```{r eval=FALSE}
a,b,c = rand(3)   # 3 random numbers in (0,1)
a < b, c*a < c*b

```
                                    

```{r eval=FALSE}
## (false,false)
```



Whenever these two commands are run, the two logical values should be identical, even though the specific values of `a`, `b`, and `c` will vary.

The restriction that $c > 0$ is needed. For example, if $c = -1$, then we have $a < b$ if and only if $-a > -b$. That is the inequality is "flipped."

```{r eval=FALSE}
a,b = rand(2)
a < b, -a > -b

```
                                    

```{r eval=FALSE}
## (false,false)
```



Again, whenever this is run, the two logical values should be the same. The values $a$ and $-a$ are the same distance from $0$, but on opposite sides. Hence if $0 < a < b$, then $b$ is farther from $0$ than $a$, so $-b$ will be farther from $0$ than $-a$, which in this case says $-b < -a$, as expected.

Finally, we have the case of division. The relation of $x$ and $1/x$ (for $x > 0$) is that the farther $x$ is from $0$, the closer $1/x$ is to $0$. So large values of $x$ make small values of $1/x$. This leads to this fact for $a,b > 0$: $a < b$ if and only if $1/a > 1/b$.

We can check with random values again:

```{r eval=FALSE}
a,b = rand(2)
a < b, 1/a > 1/b

```
                                    

```{r eval=FALSE}
## (true,true)
```



In summary we verified numerically that the following hold:

  * `a < b` if and only if `a + c < b + c` for all finite `a`, `b`, and `c`.
  * `a < b` if and only if `c*a < c*b` for all finite `a` and `b`, and finite, positive `c`.
  * `a < b` if and only if `-a > -b` for all finite `a` and `b`.
  * `a < b` if and only if `1/a > 1/b` for all finite, positive `a` and `b`.

## Some examples

We now show some inequalities highlighted on this  [Wikipedia](http://en.wikipedia.org/wiki/Inequality_%28mathematics%29) page.

Numerically investigate the fact $\exp(x) \geq 1 + x$ by showing it is true for three different values of $x$. We pick $x=-1$, $0$, and $1$:

```{r eval=FALSE}
x = -1; exp(x) >= 1 + x
x =  0; exp(x) >= 1 + x
x =  1; exp(x) >= 1 + x

```
                                    

```{r eval=FALSE}
## true
```



Now, let's investigate that for any distinct real numbers, $a$ and $b$ that
$$~
\frac{e^b - e^a}{b - a} > e^{(a+b)/2}
~$$

For this, we use `rand(2)` to generate two random numbers in $(0,1)$:

```{r eval=FALSE}
a, b = rand(2)
(exp(b) - exp(a)) / (b-a) > exp((a+b)/2)

```
                                    

```{r eval=FALSE}
## true
```



This should evaluate to `true` for any random choice of `a` and `b` returned by `rand(2)`.

Finally, let's investigate the fact that the harmonic mean, $2/(1/a + 1/b)$ is less than or equal to the geometric mean, $\sqrt{ab}$, which is less than or equal to the quadratic mean, $\sqrt{a^2 + b^2}/\sqrt{2}$, using two randomly chosen values:

```{r eval=FALSE}
a, b = rand(2)
h = 2 / (1/a + 1/b)
g = (a * b) ^ (1 / 2)
q = sqrt((a^2 + b^2) / 2)
h <= g, g <= q

```
                                    

```{r eval=FALSE}
## (true,true)
```



## Chaining, combining expressions: absolute values

The absolute value notation can be defined through cases:
$$~
\lvert x\rvert = \begin{cases}
x & x \geq 0\\
-x & \text{otherwise}.
\end{cases}
~$$

The interpretation of $\lvert x\rvert$, as the distance on the number line of $x$ from $0$, means that many relationships are naturally expressed in terms of absolute values. For example, a simple shift: $\lvert x -c\rvert$ is related to the distance $x$ is from the number $c$. As common as they are, the concept can still be confusing when inequalities are involved.

For example, the expression $\lvert x - 5\rvert < 7$ has solutions which are all values of $x$ within $7$ units of $5$. This would be the values $-2 < x < 12$. If this isn't immediately intuited, then formally $\lvert x - 5 \rvert < 7$ is a compact representation of a chain of inequalities: $-7 < x - 5 < 7$. (Which is really two combined inequalities: $-7<x-5$ *and* $x-5<7$.) We can "add" 5 to each side to get $-2 < x < 12$, using the fact that adding by a finite number does not change the inequality sign.

Julia's precedence for logical expressions, allows such statements to mirror the mathematical notation:

```{r eval=FALSE}
x = 18
abs(x - 5) < 7

```
                                    

```{r eval=FALSE}
## false
```



This is to be expected, but we could also have written:

```{r eval=FALSE}
-7 < x - 5 < 7

```
                                    

```{r eval=FALSE}
## false
```



Read aloud this would be "minus 7 is less than x minus 5 **and** x minus 5 is less than 7".

The "and" equations can be combined as above with a natural notation. However, an equation like $\lvert x - 5\rvert > 7$ would emphasize an **or** and be "x minus 5 less than minus 7 **or** x minus 5 greater than 7". Expressing this requires some new notation.

The *boolean operators* `&` and `|` implement "and" and "or." Thus we could write $-7 < x-5 < 7$ as

```{r eval=FALSE}
(-7 < x - 5) & (x - 5 < 7)

```
                                    

```{r eval=FALSE}
## false
```



and could write $\lvert x-5\rvert > 7$ as

```{r eval=FALSE}
(x - 5 < -7) | (x - 5 > 7)

```
                                    

```{r eval=FALSE}
## true
```



(The first expression is false for $x-18$ and the second expression true.)
<div class="alert alert-success" role="alert">
<div class="markdown"><p>The <a href="http://julia.readthedocs.org/en/latest/manual/control-flow/#man-short-circuit-evaluation">short circuit operators</a> are <code>&amp;&amp;</code> and <code>||</code>. For simple Boolean values, they perform a related task, though have a more general usage.</p>
</div> 
</div>

##### Example

One of [DeMorgan's Laws](http://en.wikipedia.org/wiki/De_Morgan%27s_laws) states that "not (A and B)" is the same as "(not A) or (not B)". This is a kind of distributive law for "not", but note how the "and" changes to "or". We can verify this law systematically. For example, the following shows it true for 1 of the 4 possible cases for the pair `A`, `B` to take:

```{r eval=FALSE}
A,B = true, false  ## also true, true; false, true; and false, false
!(A & B) == !A | !B

```
                                    

```{r eval=FALSE}
## true
```



## Precedence

The question of when parentheses are needed and when they are not is answered by the [precedence](http://julia.readthedocs.org/en/latest/manual/mathematical-operations/#operator-precedence)  rules implemented. Earlier, we wrote

```{r eval=FALSE}
(x - 5 < -7) | (x - 5 > 7)

```
                                    

```{r eval=FALSE}
## true
```



To represent $\lvert x-5\rvert > 7$. Were the parentheses necessary? Let's just check. 

```{r eval=FALSE}
x - 5 < -7 | x - 5 > 7

```
                                    

```{r eval=FALSE}
## false
```



So yes, they were. The precedence rules perform `|` before `<` or `>`, so without the extra pair of parentheses, we would have

```{r eval=FALSE}
(x - 5 < ( (-7 | x) - 5)) > 7

```
                                    

```{r eval=FALSE}
## false
```



which is not what is desired at all. (The value of `-7 | x` is `-5` – as `|` does something completely different when the two arguments are not boolean.)

A thorough understanding of the precedence rules can help eliminate unnecessary parentheses, but in most cases it is easier just to put them in.

## Arithmetic with

For convenience, basic arithmetic can be performed with Boolean values, `false` becomes $0$ and true $1$. For example, both these expressions make sense:

```{r eval=FALSE}
true + true + false, false * 1000

```
                                    

```{r eval=FALSE}
## (2,0)
```



The first example shows a common means used to count the number of `true` values in a collection of Boolean values – just add them.

This can be cleverly exploited. For example, the following expression returns `x` when it is positive and $0$ otherwise:

```{r eval=FALSE}
(x > 0) * x

```
                                    

```{r eval=FALSE}
## 18
```



There is a built in function, `max` that can be used for this: `max(0, x)`.

This expression returns `x` if it is between $-10$ and $10$ and otherwise $-10$ or $10$ depending on whether $x$ is negative or positive.

```{r eval=FALSE}
(x < -10)*(-10) + (x >= -10)*(x < 10) * x + (x>=10)*10

```
                                    

```{r eval=FALSE}
## 10
```



The `clamp(x, a, b)` performs this task more generally, and is used as in `clamp(x, -10, 10)`.

## Questions

###### Question

Is `e^pi` or `pi^e` greater?

----

<form name="WeaveQuestion" data-id="lcFa9Ukj" data-controltype="radio">
<div class="form-group ">

<div  class="radio"> 
<label>
<input type="radio" name="radio_lcFa9Ukj" value="1"><div class="markdown"><p><code>e^pi</code> is equal to <code>pi^e</code></p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_lcFa9Ukj" value="2"><div class="markdown"><p><code>e^pi</code> is greater than <code>pi^e</code></p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_lcFa9Ukj" value="3"><div class="markdown"><p><code>e^pi</code> is less than <code>pi^e</code></p>
</div>
</label>
</div>



<div id="lcFa9Ukj_message"></div>
</div>
</form>
<script text="text/javascript">
$("input:radio[name='radio_lcFa9Ukj']").on("change", function() {
correct = this.value == 2;

if(correct) {
$("#lcFa9Ukj_message").html("<div class='alert alert-success'><span class='glyphicon glyphicon-thumbs-up'>&nbsp;Correct</span></div>");
} else {
$("#lcFa9Ukj_message").html("<div class='alert alert-warning'><span class='glyphicon glyphicon-thumbs-down'>&nbsp;Incorrect</span></div>");
}
});

</script>

----



###### Question

Is sin(1000) positive?

----

<form name="WeaveQuestion" data-id="JBsUsiVE" data-controltype="radio">
<div class="form-group ">

<div  class="radio"> 
<label>
<input type="radio" name="radio_JBsUsiVE" value="1"><div class="markdown"><p>Yes</p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_JBsUsiVE" value="2"><div class="markdown"><p>No</p>
</div>
</label>
</div>



<div id="JBsUsiVE_message"></div>
</div>
</form>
<script text="text/javascript">
$("input:radio[name='radio_JBsUsiVE']").on("change", function() {
correct = this.value == 1;

if(correct) {
$("#JBsUsiVE_message").html("<div class='alert alert-success'><span class='glyphicon glyphicon-thumbs-up'>&nbsp;Correct</span></div>");
} else {
$("#JBsUsiVE_message").html("<div class='alert alert-warning'><span class='glyphicon glyphicon-thumbs-down'>&nbsp;Incorrect</span></div>");
}
});

</script>

----



###### Question

Suppose you know $0 < a < b$. What can you say about the relationship between $-1/a$ and $-1/b$?

----

<form name="WeaveQuestion" data-id="8B6pglIx" data-controltype="radio">
<div class="form-group ">

<div  class="radio"> 
<label>
<input type="radio" name="radio_8B6pglIx" value="1"><div class="markdown">$-1/a > -1/b$
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_8B6pglIx" value="2"><div class="markdown">$-1/a \geq -1/b$
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_8B6pglIx" value="3"><div class="markdown">$-1/a < -1/b$
</div>
</label>
</div>



<div id="8B6pglIx_message"></div>
</div>
</form>
<script text="text/javascript">
$("input:radio[name='radio_8B6pglIx']").on("change", function() {
correct = this.value == 2;

if(correct) {
$("#8B6pglIx_message").html("<div class='alert alert-success'><span class='glyphicon glyphicon-thumbs-up'>&nbsp;Correct</span></div>");
} else {
$("#8B6pglIx_message").html("<div class='alert alert-warning'><span class='glyphicon glyphicon-thumbs-down'>&nbsp;Incorrect</span></div>");
}
});

</script>

----



###### Suppose you know $a < 0 < b$, is it true that $1/a > 1/b$?

----

<form name="WeaveQuestion" data-id="wQCd3XTz" data-controltype="radio">
<div class="form-group ">

<div  class="radio"> 
<label>
<input type="radio" name="radio_wQCd3XTz" value="1"><div class="markdown"><p>It is never true, as $1/a$ is negative and $1/b$ is positive</p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_wQCd3XTz" value="2"><div class="markdown"><p>It can sometimes be true, though not always.</p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_wQCd3XTz" value="3"><div class="markdown"><p>Yes, it is always true.</p>
</div>
</label>
</div>



<div id="wQCd3XTz_message"></div>
</div>
</form>
<script text="text/javascript">
$("input:radio[name='radio_wQCd3XTz']").on("change", function() {
correct = this.value == 1;

if(correct) {
$("#wQCd3XTz_message").html("<div class='alert alert-success'><span class='glyphicon glyphicon-thumbs-up'>&nbsp;Correct</span></div>");
} else {
$("#wQCd3XTz_message").html("<div class='alert alert-warning'><span class='glyphicon glyphicon-thumbs-down'>&nbsp;Incorrect</span></div>");
}
});

</script>

----



###### Question

The `airy` [function](http://en.wikipedia.org/wiki/Airy_function) is a special function named after a British Astronomer who realized the function's value in his studies of the rainbow. It is known that it is always positive for $x > 0$, though not so for negative values of $x$. Which of these indicates the first negative value : `airy(-1) < 0`, `airy(-2) < 0`, ... `airy(-5) < 0`?

----

<form name="WeaveQuestion" data-id="AhffdyBs" data-controltype="radio">
<div class="form-group ">

<div  class="radio"> 
<label>
<input type="radio" name="radio_AhffdyBs" value="1"><div class="markdown"><p><code>airy&#40;-1&#41; &lt; 0</code></p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_AhffdyBs" value="2"><div class="markdown"><p><code>airy&#40;-2&#41; &lt; 0</code></p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_AhffdyBs" value="3"><div class="markdown"><p><code>airy&#40;-3&#41; &lt; 0</code></p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_AhffdyBs" value="4"><div class="markdown"><p><code>airy&#40;-4&#41; &lt; 0</code></p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_AhffdyBs" value="5"><div class="markdown"><p><code>airy&#40;-5&#41; &lt; 0</code></p>
</div>
</label>
</div>



<div id="AhffdyBs_message"></div>
</div>
</form>
<script text="text/javascript">
$("input:radio[name='radio_AhffdyBs']").on("change", function() {
correct = this.value == 3;

if(correct) {
$("#AhffdyBs_message").html("<div class='alert alert-success'><span class='glyphicon glyphicon-thumbs-up'>&nbsp;Correct</span></div>");
} else {
$("#AhffdyBs_message").html("<div class='alert alert-warning'><span class='glyphicon glyphicon-thumbs-down'>&nbsp;Incorrect</span></div>");
}
});

</script>

----



###### Question

By trying three different values of $x > 0$ which of these could possibly be always true:

----

<form name="WeaveQuestion" data-id="SrAYjbkB" data-controltype="radio">
<div class="form-group ">

<div  class="radio"> 
<label>
<input type="radio" name="radio_SrAYjbkB" value="1"><div class="markdown"><p><code>x^x &lt;&#61; &#40;1/e&#41;^&#40;1/e&#41;</code></p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_SrAYjbkB" value="2"><div class="markdown"><p><code>x^x &gt;&#61; &#40;1/e&#41;^&#40;1/e&#41;</code></p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_SrAYjbkB" value="3"><div class="markdown"><p><code>x^x &#61;&#61; &#40;1/e&#41;^&#40;1/e&#41;</code></p>
</div>
</label>
</div>



<div id="SrAYjbkB_message"></div>
</div>
</form>
<script text="text/javascript">
$("input:radio[name='radio_SrAYjbkB']").on("change", function() {
correct = this.value == 2;

if(correct) {
$("#SrAYjbkB_message").html("<div class='alert alert-success'><span class='glyphicon glyphicon-thumbs-up'>&nbsp;Correct</span></div>");
} else {
$("#SrAYjbkB_message").html("<div class='alert alert-warning'><span class='glyphicon glyphicon-thumbs-down'>&nbsp;Incorrect</span></div>");
}
});

</script>

----



###### Question

Student logic says $(x+y)^p = x^p + y^p$. Of course, this isn't correct for all $p$ and $x$. By trying a few points, which is true when $x,y > 0$ and $0 < p < 1$:

----

<form name="WeaveQuestion" data-id="OUfJhty3" data-controltype="radio">
<div class="form-group ">

<div  class="radio"> 
<label>
<input type="radio" name="radio_OUfJhty3" value="1"><div class="markdown"><p><code>&#40;x&#43;y&#41;^p &gt; x^p &#43; y^p</code></p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_OUfJhty3" value="2"><div class="markdown"><p><code>&#40;x&#43;y&#41;^p &#61;&#61; x^p &#43; y^p</code></p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_OUfJhty3" value="3"><div class="markdown"><p><code>&#40;x&#43;y&#41;^p &lt; x^p &#43; y^p</code></p>
</div>
</label>
</div>



<div id="OUfJhty3_message"></div>
</div>
</form>
<script text="text/javascript">
$("input:radio[name='radio_OUfJhty3']").on("change", function() {
correct = this.value == 3;

if(correct) {
$("#OUfJhty3_message").html("<div class='alert alert-success'><span class='glyphicon glyphicon-thumbs-up'>&nbsp;Correct</span></div>");
} else {
$("#OUfJhty3_message").html("<div class='alert alert-warning'><span class='glyphicon glyphicon-thumbs-down'>&nbsp;Incorrect</span></div>");
}
});

</script>

----



###### Question

According to Wikipedia, one of the following inequalities is always true for $a, b > 0$ (as proved by I. Ilani in JSTOR,AMM,Vol.97,No.1,1990). Which one?

----

<form name="WeaveQuestion" data-id="BjWJgref" data-controltype="radio">
<div class="form-group ">

<div  class="radio"> 
<label>
<input type="radio" name="radio_BjWJgref" value="1"><div class="markdown"><p><code>a^a &#43; b^b &gt;&#61; a^b &#43; b^a</code></p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_BjWJgref" value="2"><div class="markdown"><p><code>a^b &#43; b^a &lt;&#61; 1</code></p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_BjWJgref" value="3"><div class="markdown"><p><code>a^a &#43; b^b &lt;&#61; a^b &#43; b^a</code></p>
</div>
</label>
</div>



<div id="BjWJgref_message"></div>
</div>
</form>
<script text="text/javascript">
$("input:radio[name='radio_BjWJgref']").on("change", function() {
correct = this.value == 1;

if(correct) {
$("#BjWJgref_message").html("<div class='alert alert-success'><span class='glyphicon glyphicon-thumbs-up'>&nbsp;Correct</span></div>");
} else {
$("#BjWJgref_message").html("<div class='alert alert-warning'><span class='glyphicon glyphicon-thumbs-down'>&nbsp;Incorrect</span></div>");
}
});

</script>

----



###### Question

Is $3$ in the set $\lvert x - 2\rvert < 1/2$?

----

<form name="WeaveQuestion" data-id="w9sjfGgJ" data-controltype="radio">
<div class="form-group ">

<div  class="radio"> 
<label>
<input type="radio" name="radio_w9sjfGgJ" value="1"><div class="markdown"><p>Yes</p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_w9sjfGgJ" value="2"><div class="markdown"><p>No</p>
</div>
</label>
</div>



<div id="w9sjfGgJ_message"></div>
</div>
</form>
<script text="text/javascript">
$("input:radio[name='radio_w9sjfGgJ']").on("change", function() {
correct = this.value == 2;

if(correct) {
$("#w9sjfGgJ_message").html("<div class='alert alert-success'><span class='glyphicon glyphicon-thumbs-up'>&nbsp;Correct</span></div>");
} else {
$("#w9sjfGgJ_message").html("<div class='alert alert-warning'><span class='glyphicon glyphicon-thumbs-down'>&nbsp;Incorrect</span></div>");
}
});

</script>

----



###### Question

Which of the following is equivalent to $\lvert x - a\rvert > b$:

----

<form name="WeaveQuestion" data-id="goTRFs41" data-controltype="radio">
<div class="form-group ">

<div  class="radio"> 
<label>
<input type="radio" name="radio_goTRFs41" value="1"><div class="markdown"><p>$ x - a < -b \text{ or } x - a > b$</p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_goTRFs41" value="2"><div class="markdown">$-b < x - a < b$
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_goTRFs41" value="3"><div class="markdown"><p>$ -b < x-a \text{ and } x - a < b$</p>
</div>
</label>
</div>



<div id="goTRFs41_message"></div>
</div>
</form>
<script text="text/javascript">
$("input:radio[name='radio_goTRFs41']").on("change", function() {
correct = this.value == 1;

if(correct) {
$("#goTRFs41_message").html("<div class='alert alert-success'><span class='glyphicon glyphicon-thumbs-up'>&nbsp;Correct</span></div>");
} else {
$("#goTRFs41_message").html("<div class='alert alert-warning'><span class='glyphicon glyphicon-thumbs-down'>&nbsp;Incorrect</span></div>");
}
});

</script>

----



###### Question

If $\lvert x - \pi\rvert < 1/10$ is $\lvert \sin(x) - \sin(\pi)\rvert < 1/10$?

Guess an answer based on a few runs of
<pre class="sourceCode julia">x = pi + 0.2 * rand()
abs(x - pi) < 1/10, abs(sin(x) - sin(pi)) < 1/10
</pre>
----

<form name="WeaveQuestion" data-id="oxJOhRMT" data-controltype="radio">
<div class="form-group ">

<div  class="radio inline"> 
<label>
<input type="radio" name="radio_oxJOhRMT" value="1"><div class="markdown"><p>true</p>
</div>
</label>
</div>

<div  class="radio inline"> 
<label>
<input type="radio" name="radio_oxJOhRMT" value="2"><div class="markdown"><p>false</p>
</div>
</label>
</div>



<div id="oxJOhRMT_message"></div>
</div>
</form>
<script text="text/javascript">
$("input:radio[name='radio_oxJOhRMT']").on("change", function() {
correct = this.value == 1;

if(correct) {
$("#oxJOhRMT_message").html("<div class='alert alert-success'><span class='glyphicon glyphicon-thumbs-up'>&nbsp;Correct</span></div>");
} else {
$("#oxJOhRMT_message").html("<div class='alert alert-warning'><span class='glyphicon glyphicon-thumbs-down'>&nbsp;Incorrect</span></div>");
}
});

</script>

----



###### Question

Does `12` satisfy $\lvert x - 3\rvert + \lvert x-9\rvert > 12$?

----

<form name="WeaveQuestion" data-id="Em8gzP0F" data-controltype="radio">
<div class="form-group ">

<div  class="radio"> 
<label>
<input type="radio" name="radio_Em8gzP0F" value="1"><div class="markdown"><p>Yes</p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_Em8gzP0F" value="2"><div class="markdown"><p>No</p>
</div>
</label>
</div>



<div id="Em8gzP0F_message"></div>
</div>
</form>
<script text="text/javascript">
$("input:radio[name='radio_Em8gzP0F']").on("change", function() {
correct = this.value == 2;

if(correct) {
$("#Em8gzP0F_message").html("<div class='alert alert-success'><span class='glyphicon glyphicon-thumbs-up'>&nbsp;Correct</span></div>");
} else {
$("#Em8gzP0F_message").html("<div class='alert alert-warning'><span class='glyphicon glyphicon-thumbs-down'>&nbsp;Incorrect</span></div>");
}
});

</script>

----



###### Question

Which of these will show DeMorgan's law holds when both values are `false`:

----

<form name="WeaveQuestion" data-id="ipPKLiDZ" data-controltype="radio">
<div class="form-group ">

<div  class="radio"> 
<label>
<input type="radio" name="radio_ipPKLiDZ" value="1"><div class="markdown"><p><code>&#33;&#40;false &amp; false&#41; &#61;&#61; &#33;false &amp; &#33;false</code></p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_ipPKLiDZ" value="2"><div class="markdown"><p><code>&#33;&#40;false &amp; false&#41; &#61;&#61; false &#33;&amp; false</code></p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_ipPKLiDZ" value="3"><div class="markdown"><p><code>&#33;&#40;false &amp; false&#41; &#61;&#61; &#33;false | &#33;false</code></p>
</div>
</label>
</div>



<div id="ipPKLiDZ_message"></div>
</div>
</form>
<script text="text/javascript">
$("input:radio[name='radio_ipPKLiDZ']").on("change", function() {
correct = this.value == 3;

if(correct) {
$("#ipPKLiDZ_message").html("<div class='alert alert-success'><span class='glyphicon glyphicon-thumbs-up'>&nbsp;Correct</span></div>");
} else {
$("#ipPKLiDZ_message").html("<div class='alert alert-warning'><span class='glyphicon glyphicon-thumbs-down'>&nbsp;Incorrect</span></div>");
}
});

</script>

----



###### Question

For floating point numbers there are two special values `Inf` and `NaN`. For which of these is the answer always `false`:

----

<form name="WeaveQuestion" data-id="Dn67fUIv" data-controltype="radio">
<div class="form-group ">

<div  class="radio"> 
<label>
<input type="radio" name="radio_Dn67fUIv" value="1"><div class="markdown"><p><code>Inf &lt; 3.0</code> and <code>3.0 &lt;&#61; Inf</code></p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_Dn67fUIv" value="2"><div class="markdown"><p><code>NaN &lt; 3.0</code> and <code>3.0 &lt;&#61; NaN</code></p>
</div>
</label>
</div>



<div id="Dn67fUIv_message"></div>
</div>
</form>
<script text="text/javascript">
$("input:radio[name='radio_Dn67fUIv']").on("change", function() {
correct = this.value == 2;

if(correct) {
$("#Dn67fUIv_message").html("<div class='alert alert-success'><span class='glyphicon glyphicon-thumbs-up'>&nbsp;Correct</span></div>");
} else {
$("#Dn67fUIv_message").html("<div class='alert alert-warning'><span class='glyphicon glyphicon-thumbs-down'>&nbsp;Incorrect</span></div>");
}
});

</script>

----



###### Question

The IEEE 754 standard  is about floating point numbers, for which there are the special values `Inf`, `-Inf`, `NaN`, and (surprisingly `-0`. Here are 4 facts that seem reasonable:

  * Positive zero is equal but not greater than negative zero.
  * `Inf` is equal to itself and greater than everything else except `NaN`.
  * `-Inf` is equal to itself and less then everything else except `NaN`.
  * `NaN` is not equal to, not less than, and not greater than anything, including itself.

Do all four seem to be the case within `Julia`? Find your answer by trial and error.

----

<form name="WeaveQuestion" data-id="UpJYMEZs" data-controltype="radio">
<div class="form-group ">

<div  class="radio"> 
<label>
<input type="radio" name="radio_UpJYMEZs" value="1"><div class="markdown"><p>Yes</p>
</div>
</label>
</div>

<div  class="radio"> 
<label>
<input type="radio" name="radio_UpJYMEZs" value="2"><div class="markdown"><p>No</p>
</div>
</label>
</div>



<div id="UpJYMEZs_message"></div>
</div>
</form>
<script text="text/javascript">
$("input:radio[name='radio_UpJYMEZs']").on("change", function() {
correct = this.value == 1;

if(correct) {
$("#UpJYMEZs_message").html("<div class='alert alert-success'><span class='glyphicon glyphicon-thumbs-up'>&nbsp;Correct</span></div>");
} else {
$("#UpJYMEZs_message").html("<div class='alert alert-warning'><span class='glyphicon glyphicon-thumbs-down'>&nbsp;Incorrect</span></div>");
}
});

</script>

----



